Using simulation to optimize the sampling design of fisheries-independent trawl surveys

### 2018-07-09 analysis/paper/sim_survey_paper.Rmd
Start parameter setting table

---
title: Using simulation to optimize the sampling design of fisheries-independent trawl surveys
author: "Paul M. Regular, Fran Mowbray, et al.?"
date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
output: 
  word_document:
    reference_docx: template.docx
bibliography: references.bib
csl: ices-journal-of-marine-science.csl
---

@@

$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
Here, numbers-at-age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e. recruits) are filled using a random walk, $log(N_{1,y}) = log(N_{1,y-1}) + \epsilon_{y-1}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(z) + \delta_{a,y}$. The error around the recruitment process was simulated using the normal distribution, $\epsilon_{y} \sim N(0, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$. The covariance across ages and years is controled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, age}$ and $\varphi_{\delta, year}$, respectively). This structure allows for autocorrelation in process errors across ages and years (i.e. total mortality can be made to be more similar for fish that are closer together in age and time). Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.

@@

The next step in the simulation is to distribute the abundance-at-age matrix simulated using the cohort model throughout a spatial field. Here, a grid of $s$ locations was generated with a resolution of $r$ and a depth gradient. The depth gradient was set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$. The simulated population is distributed through the grid by simulating spatial-temporal noise controled by a parabolic relationship with depth and covariance between ages, years and space

@@

Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or disperson of abundance around the mean. This helps simulate depth preferences of praticular fish. Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years. Spatial correlations was controled by a smoothing ($\lambda$) and scaling parameter ($\kappa$) [@Blangiardo2015] and correlation across ages and years was controled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively. The overall variance of the process was controled by $\sigma_{\xi}^2$. In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering. Note that $\eta_{a,y,s}$ is forced to sum to one in each year to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model. Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid. A more detailed description of the space-age-year covariance is included in Appendix 1.

@@

The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated using formula described above. First, the grid is divided into $N_{div}$ divisions and $N_{strat}$ depth-based strata. The area of each strata is calculated and this is used to define the sets to allocate to each strata under a praticular set density, $D_{sets}$. The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:

$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A}{r^2} q_a )$$
Where $n_{a,y,s}$ is the number of fish sampled by a set, $A$ indicates the area covered by the trawl, $r$ is the resolution of the grid cells, and $q_a$ is the catchability coefficient of each age. Here, catchabilities were defined using a logistic curve controled by a steepness, $k$, and midpoint parameter, $x_0$. The lengths of the fish sampled by the set are then simulated using the original von Bertalanffy growth curve [@Cailliet2006]:

$$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$
Where $l_{\infty}$ is the mean asympotic lengths, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$. Depending on the number of fish caught, sub-sampling is then conducted. Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division.

@@

>>>>

---
title: Using simulation to optimize the sampling design of fisheries-independent trawl surveys
author: "Paul M. Regular, Fran Mowbray, et al.?"
date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
bibliography: references.bib
csl: ices-journal-of-marine-science.csl
---

@@

$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
Here, numbers-at-age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e. recruits) vary around a baseline value, $log(N_{1,y}) = log(\mu_r) + \epsilon_{y}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(\mu_Z) + \delta_{a,y}$. The error around the recruitment process was set to follow a random walk, $\epsilon_{y} \sim N(\epsilon_{y - 1}, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$. The covariance across ages and years is controlled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, age}$ and $\varphi_{\delta, year}$, respectively). This structure allows for autocorrelation in process errors across ages and years (i.e. total mortality can be made to be more similar for fish that are closer together in age and time). Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.

@@

The next step in the simulation is to distribute the abundance-at-age matrix simulated using the cohort model throughout a spatial field. Here, a grid of $s$ locations was generated with a resolution of $r$ and a depth gradient. The depth gradient was set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$. The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space

@@

Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean. This helps simulate depth preferences of particular fish. Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years. Spatial correlations was controlled by a smoothing ($\lambda$) and scaling parameter ($\kappa$) [@Blangiardo2015] and correlation across ages and years was controlled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively. The overall variance of the process was controlled by $\sigma_{\xi}^2$. In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering. Note that $\eta_{a,y,s}$ is forced to sum to one in each year to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model. Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid. A more detailed description of the space-age-year covariance is included in Appendix 1.

@@

The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated using formula described above. First, the grid is divided into $N_{div}$ divisions and $N_{strat}$ depth-based strata. The area of each strata is calculated and this is used to define the sets to allocate to each strata under a particular set density, $D_{sets}$. The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:

$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A}{r^2} q_a )$$
Where $n_{a,y,s}$ is the number of fish sampled by a set, $A$ indicates the area covered by the trawl, $r$ is the resolution of the grid cells, and $q_a$ is the catchability coefficient of each age. Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$. The lengths of the fish sampled by the set are then simulated using the original von Bertalanffy growth curve [@Cailliet2006]:

$$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$
Where $l_{\infty}$ is the mean asymptotic lengths, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$. Depending on the number of fish caught, sub-sampling is then conducted. Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division.

@@

Ideally the parameter settings used here would be based on model fits to real data, however, conditioning the operating model described above was outside the scope of this project. Moreover, many of the parameters described above may not be identifiable or it may be difficult to obtain un-biased estimates [e.g. bias introduced by length-stratified sampling complicate the fitting of growth models; @Morgan1997]. As such, parameters were informally selected to roughly correspond to observations typically obtained by the Fisheries and Oceans Canada survey of cod in NAFO division 3Ps, Newfoundland. Though this approach is far from exact, the goal was to approximate the life-history and distribution of an actual population to simulate an arena for testing the consequences of surveys with different sampling protocol. Parameter settings and descriptions used in this case study are presented in Table \@ref(tab:test).

# Tables

Table: (\#tab:test) Parameter descriptions, symbols and settings used in the cod case study.

**Parameter name**   **Symbol**   **Setting**
------------------   ----------   -----------
              Ages          $a$          1:20
             Years          $y$          1:20
  Mean recruitment      $\mu_r$           30M
